Quantifying Directionality and Innovation Output in Sweden's Transition to a Forest-Based Bioeconomy

### 2024-05-01 paper.qmd
update paper

---
title: Directionality and Direction of Innovation in the Swedish Forest-Based Bioeconomy Transition 
subtitle: Draft
author: 
    name: Jonas Kreutzer
    affiliation: Department of Economic History, Lund University
date: last-modified
abstract: |
    The transition to a bioeconomy is high on the agenda for research, industry and policy. Beyond conceptual uncertainties of what is meant by a bioeconomy, three different visions have emerged, differing in their aims and objectives, drivers and mediators of innovation and relation to the biosphere. The use of detailed innovation data from a Swedish innovation database, SWINNO, allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system. Qualitative coding of the underlying source material associates innovation from 1970 to 2010 with visions of a bioresource, biotechnology and / or bioecology bioeconomy. Preliminary results suggest that the innovation output of the forest based bioeconomy has decreased relative to the total output. Of the three visions, the bioresource vision has been the most dominant, especially in the 1970s and after 2005. At the same time, innovation which fit into multiple visions exist, suggesting that the opportunity space for future development is not yet locked in.

csl: https://zotero.org/styles/apa?source=1
bibliography: [swinno-shades-green.bib]
link-citations: true
link-bibliography: true
colorlinks: true
number-sections: true

execute: 
  warning: false
  error: false
  echo: false

format:
    pdf:
        documentclass: scrartcl
        papersize: A4
        linestretch: 1.25
---

@@

import pandas as pd
import numpy as np
from IPython.display import display, Markdown
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import seaborn as sns

from src.figures import *
from src.utils import get_project_root, format_table
from src.swinno_helpers import connect_swinno_db

ROOT = get_project_root()
conn = connect_swinno_db()

@@

sns.set_theme(style="white")

@@

codes_df = codes_df.set_index("Code")
```

```{python}
# | label: Query bioeconomy

bioeconomy = pd.read_sql(
    """ 
    SELECT distinct sinno_id from (SELECT
    i.sinno_id,
    i.innovation_name_in_swedish AS name,
    i.literal_name_of_innovating_firm_sinnos_common_name_sinno_id AS firm,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id AS collaborator_0,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id1 AS collaborator_1,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id2 AS collaborator_2,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id3 AS collaborator_3,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id4 AS collaborator_4,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id5 AS collaborator_5,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id6 AS collaborator_6,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id7 AS collaborator_7,
    i.literal_name_of_collaborator_sinnos_common_name_sinno_id8 AS collaborator_8,
    i.description_in_swedish AS description,
    i.year_of_commercialization AS year,
    us.use_sector
    FROM
    innovation i
    JOIN use_sectors us ON i.sinno_id = us.sinno_id
    WHERE
    (
        us.use_sector LIKE '02%'
        OR us.use_sector LIKE '20%'
        OR us.use_sector LIKE '21%'
        OR us.use_sector LIKE '36%'
        OR product_code LIKE '02%'
        OR product_code LIKE '20%'
        OR product_code LIKE '21%'
        OR product_code LIKE '36%'
    )
    OR (
        description LIKE '%virke%'
        OR description LIKE '%cellulos%'
        OR description LIKE '%lignin%'
        OR description LIKE '%spån%'
        OR description LIKE '%bark%'
        OR description LIKE '%levulinsyra%'
        OR description LIKE '%furfural%'
        OR description LIKE '%svarttjära%'
        OR description LIKE '%svartlut%'
        OR description LIKE '%växtbas%'
        OR description LIKE '%ved%'
        OR description LIKE '%trä%'
        OR description LIKE '%skog%'
        OR description LIKE '%biobränsle%'
        OR description LIKE '%biologisk%'
        OR description LIKE '%nedbrytbar%'
        OR description LIKE '%papper%'
        OR description LIKE '%pappret%'
        OR description LIKE '%karton%'
        OR description LIKE '%tencel%'
    )
    AND i.sinno_id not in (select sinno_id from categorization_notes
    where notes not like "%not forest%"));
""",
    conn,
)

swinno["bioeconomy"] = swinno["sinno_id"].isin(bioeconomy["sinno_id"].unique())

```

@@

<!-- Paragraphs 1-2: Motivation. After reading these paragraphs a reader in any field of economics should believe that if you answer your research question your paper will make an important contribution. -->
Discussions about transitions to a bioeconomy are a hot topic of research and policy at the moment.
Different visions of the endgoal exists, stressing needs for ecological, resource or technological aims.
Actors in the transition stress the need for innovation to transition to a bioeconomy.
Yet, the actual efforts undertaken to realize these visions remain underexplored.

<!-- • Paragraphs 3-4: Challenges. These paragraphs explain why your research question has not already been answered, i.e., what are the central challenges a researcher must tackle to answer this question. -->
Rigorously operationalizing directionality is challening due to a plurality of possible development paths [@stirling2011PluralisingProgressIntegrative].
Stirling argues that innovation is better understood as a vector of possibilities rather than a scalar decision between clearly defined alternatives.
<!-- Lastly, Stirling (2011, 2009, 2008) is a clear voice pointing out a slightly different implication of directionality. Arguably, he introduced the concept to the field of sustainability transitions research, as part of broader critique of one-dimensional growth ideology. In line with a multidimensional view on sociotechnical change, he stresses that innovation is better viewed as “a vector, rather than just a scalarquantity” (Stirling, 2011, p. 84, 2008, p. 263). But he goes further and makes a passionate and eloquent call for a more pluralistic perspective on technological innovation.

@andersson2021OutcomesDirectionalityMorphology 111 -->
Where innovation indicators are concerned, directionality is frequently constrained to a dichotomous choice by the innovator, for example, to produce an environmental innovation or to produce a conventional innovation.
In reality, however, innovators face a much more uncertain decision with many possible paths, of which some might be exclusionary.
Additionally, these studies frequently focus on patents as proxies for innovation acitvity.
Patents, however, suffer from the flaw that they may be filed for strategic reasons for example, to prevent a competitor from accessing a technolgy.
Care should therefore be taken to infer the direction of a transition from patent fillings alone.

Previous research from transition studies and political sciences has been more succesful in illuminating the pluarlity of possible tranisiton directions.
However, the basis for this is usually confined to the analysis of discourse, which can illuminate the intent, but as the saying goes "Talk is cheap".
Without investigating what actors are willing to implement discourse analysis remains a poor measure of directionality in innovation directionality.

A central challenge in assessing the directionality of transitions lies in lack of high quality data that allows to unpack which of the intended directions are seriously pursued by innovators.

@@

```{python}
# | label: Total SWINNO Innovation
# | output: asis

@@

{{< include _decision-matrix.md >}}
: Decision matrix used to classify innovation. {#tbl-decision-matrix tbl-colwidths="[20,80]"}

@@

```{python}
# | label: tbl-vision-examples
# | tbl-cap: "List of Eco-Innovation Labels"

# classification_examples = pd.read_sql(
#     """SELECT
#   n.sinno_id,
#   i.description_in_swedish,
#   GROUP_CONCAT(c.Category, ' & ') AS Categories,
#   n.notes
# FROM
#   categorization_notes n
#   JOIN innovation i ON i.sinno_id = n.sinno_id
#   JOIN bioeconomy_visions bv
#   ON n.sinno_id = bv.sinno_id
#   JOIN classification_codes c ON bv.bioeconomy_vision = c.Code
# WHERE
#   n.notes LIKE "%good example%"
# GROUP BY
#   n.sinno_id,
#   i.description_in_swedish,
#   n.notes""",
#     conn,
# )

# classification_examples_out = (
#     classification_examples.copy()
#     .loc[:, ["Categories", "description_in_swedish"]]
#     .iloc[2:]
# )

# rowIDs = range(1, len(classification_examples_out) + 1)
# display(
#     Markdown(
#         classification_examples_out.rename(
#             columns={"description_in_swedish": "Short Description"}
#         ).to_markdown(index=rowIDs)
#     )
# )

```

### Eco-Innovation Types {#sec-eco_innovatoin_types}

@@

```{python}
# | label: Number Bioeconomy Innovation
# | output: asis

sum_bioeconomy_innovations = swinno["bioeconomy"].sum()

```

@@

```{python}
# | label: Query eco-innovation

eco_innovations = pd.read_sql(
    """
SELECT
  sinno_id,
  innovation_type as innovation_type_code,
  c.Category as innovation_type
FROM
  eco_innovations
JOIN classification_codes AS c ON c.Code = eco_innovations.innovation_type
WHERE 
innovation_type NOT IN (701, 602, 601, 501, 111, 109, 108)
;
""",
    conn,
)

eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()

```

```{python}
# | label: Calculate Bioeconomy Share

@@

bioeconomy_share = calculate_bioeconomy_share(swinno)

```

```{python}
# | label: Bioeconomy Trend Description
# | output: asis

share_max = bioeconomy_share["bioeconomy_share"].max()
share_max_year = bioeconomy_share.loc[
    bioeconomy_share["bioeconomy_share"] == share_max, "year"
].values[0]
share_last_year = bioeconomy_share["year"].max()
share_latest = bioeconomy_share.loc[
    bioeconomy_share["year"] == share_last_year, "bioeconomy_share"
].values[0]

```

```{python}
# | label: fig-bioeconomy-share
# | fig-cap: "Bioeconomy Innovation Trends in SWINNO (Uncleaned)"
# | fig-subcap:
# |   - "Bioeconomy Innovation Count"
# |   - "Bioeconomy Innovation Share in %"
# | layout-ncol: 1

plot_count(bioeconomy_share)
plot_share(bioeconomy_share)
```

@@

```{python}
# | label: Query bioeconomy visions

bioeconomy_visions = pd.read_sql(
    """
SELECT bv.sinno_id
, codes.Category AS bioeconomy_vision
, i.year_of_commercialization AS "year"
FROM bioeconomy_visions AS bv
JOIN [classification_codes] AS codes ON codes.code = bv.bioeconomy_vision
JOIN innovation AS i ON i.sinno_id = bv.sinno_id
WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
UNION
SELECT DISTINCT(ei.sinno_id)
, 'Bioecology Vision' AS bioeconomy_vision
, i.year_of_commercialization AS "i.year"
FROM eco_innovations AS ei
JOIN innovation AS i ON i.sinno_id = ei.sinno_id
WHERE ei.innovation_type IN (206, 107)
AND ei.sinno_id NOT IN (
  SELECT ei2.sinno_id
  FROM eco_innovations as ei2
  WHERE ei2.innovation_type IN (999, '000')
  );
-- adds recycling innovation to bioecology
""",
    conn,
)

@@

total_certain_bioeconomy_visions = (
    count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
    .agg("sum", numeric_only=True)
    .sort_values("count", ascending=False)
    .reset_index()
)

```

```{python}
# | label: Number of Uncertains
# | output: asis

n_uncertain_bioeconomy_visions = uncertain_bioeconomy_visions["sinno_id"].nunique()

display(
    Markdown(
        f"For the following part, innovations which do not relate to the forest are excluded. Excluded are also the {n_uncertain_bioeconomy_visions} innovations whose source text did not allow labeling without significant uncertaintiy. These innovations are planned to undergo a revision by subject matter experts at a later point."
    )
)
```

```{python}
# | label: Number of Results
# | output: asis

@@

display(
    Markdown(
        f"This leaves {n_certain_bioeconomy_visions} innovations have been classified for the period between {certain_bioeconomy_visions_min_year} and {certain_bioeconomy_visions_max_year}. With {n_most_bioeconomy_vision} innovations, the {total_certain_bioeconomy_visions.iloc[0, 0]} has the most associated innovations. Next the {total_certain_bioeconomy_visions.iloc[1, 0]} was associated with {total_certain_bioeconomy_visions.iloc[1,1]}, while {total_certain_bioeconomy_visions.iloc[2,1]} and  {total_certain_bioeconomy_visions.iloc[3,1]} innovation were classified as {total_certain_bioeconomy_visions.iloc[2,0]} and {total_certain_bioeconomy_visions.iloc[3,0]}, respectively. Note that the sum by category will not yield the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision."
    )
)

```

@@

```{python}
# | label: Count Recycling Innovation

@@

```{python}
# | label: tbl-count_bioeconomy_visions
# | tbl-cap: "Count of Bioeconomy Visions"
# | output: asis
# | include: false

@@

```{python}
# | label: fig-bioeconomy-trends
# | fig-cap: "Count of Bioeconomy Visions by Year"

plot_bv_trends(count_certain_bioeconomy_visions)

```

@@

```{python}
# | label: fig-vision-heatmap
# | fig-caption: "Sector's Contribution to Bioeconomy Visions"
# | include: false

# sectors = pd.read_sql(
#     """
# SELECT
#   bv.sinno_id as sinno_id,
#   c.Category AS vision,
#   sni_codes.label AS sector
# FROM
#   bioeconomy_visions AS bv
#   JOIN use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
#   JOIN sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
#   JOIN classification_codes AS c ON bv.bioeconomy_vision = c.Code;
# """,
#     conn,
# )

# sectors = sectors.loc[
#     sectors["sinno_id"].isin(set(certain_bioeconomy_visions["sinno_id"])), :
# ]

# sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "Vision Neutral"

# sector_counts = (
#     sectors.groupby(["sector", "vision"]).size().to_frame("count").reset_index()
# )

# short_sector_labels = {
#     "Manufacture of other transport equipment": "Transport Equipment",
#     "Manufacture of motor vehicles, trailers and semi-trailers": "Motor Vehicle Manufacturing",
#     "Manufacture of furniture; manufacturing n.e.c.": "Furniture Manufacturing",
#     "Manufacture of fabricated metal products, except machinery and equipment": "Metal Products Manufacturing",
#     "Manufacture of machinery and equipment n.e.c.": "Machinery and Equipment",
#     "Construction": "Construction",
#     "Manufacture of electrical machinery and apparatus n.e.c.": "Electrical Machinery Manufacturing",
#     "Manufacture of radio, television and communication equipment and apparatus": "Communication Equipment",
#     "Manufacture of rubber and plastic products": "Rubber and Plastic Products",
#     "Forestry, logging and related service activities": "Forestry and Logging",
#     "Manufacture of wood and of products of wood and cork, except furniture; manufacture of articles of straw and plaiting materials": "Wood and Cork Products",
#     "Agriculture, hunting and related service activities": "Agriculture and Hunting",
#     "Land transport; transport via pipelines": "Land Transport",
#     "Supporting and auxiliary transport activities; activities of travel agencies": "Transport Services",
#     "Water transport": "Water Transport",
#     "Air transport": "Air Transport",
#     "Manufacture of medical, precision and optical instruments, watches and clocks": "Precision Instruments",
#     "Recycling": "Recycling",
#     "Manufacture of basic metals": "Basic Metals Manufacturing",
#     "Public administration and defence; compulsory social security": "Public Administration",
#     "Electricity, gas, steam and hot water supply": "Utilities",
#     "Sewage and refuse disposal, sanitation and similar activities": "Waste Management",
#     "Manufacture of chemicals and chemical products": "Chemical Products",
#     "Manufacture of pulp, paper and paper products": "Paper Products",
#     "Manufacture of coke, refined petroleum products and nuclear fuel": "Petroleum Products",
#     "Collection, purification and distribution of water": "Water Services",
#     "Publishing, printing and reproduction of recorded media": "Publishing and Media",
#     "Other mining and quarrying": "Mining and Quarrying",
#     "Mining of metal ores": "Metal Ore Mining",
#     "Tanning and dressing of leather; manufacture of luggage, handbags, saddlery, harness and footwear": "Leather and Footwear",
#     "Manufacture of food products and beverages": "Food and Beverages",
#     "Manufacture of other non-metallic mineral products": "Non-metallic Products",
#     "Wholesale trade and commission trade, except of motor vehicles and motorcycles": "Wholesale Trade",
#     "Other business activities": "Business Services",
#     "Research and development": "R&D",
#     "Manufacture of textiles": "Textile Manufacturing",
#     "Financial intermediation, except insurance and pension funding": "Financial Services",
#     "Education": "Education",
#     "Health and social work": "Healthcare",
#     "Fishing, operation of fish hatcheries and fish farms; service activities incidental to fishing": "Fishing and Aquaculture",
#     "Recreational, cultural and sporting activities": "Recreation and Culture",
#     "Extraction of crude petroleum and natural gas; service activities incidental to oil and gas extraction excluding surveying": "Oil and Gas Extraction",
#     "Post and telecommunications": "Post and Telecom",
#     "Sale, maintenance and repair of motor vehicles and motorcycles; retail sale of automotive fuel": "Automotive Services",
#     "Manufacture of wearing apparel; dressing and dyeing of fur": "Apparel Manufacturing",
# }

# sector_counts = sector_counts.sort_values("count", ascending=False)

# sector_counts["sector"] = sector_counts["sector"].map(short_sector_labels)

# heatmap_data = prepare_heatmap_data(sector_counts, "sector", "vision", "count")

# plot_heatmap(heatmap_data)

```

@@

A noteworthy feature of innovation belonging to No Bioresource Vision is that this category has by far the most Non-Eco innovations.
For the sustainability transition of the forest bioeconomy this could imply that having some Vision of a bioeconomy creates more eco-innovation than pursuing other innovation paths.

@@

@@

```{python}
# | label: fig-eco-innovation-bioeconomy-vision
# | fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
p = plot_eco_innovation_by_bioeconomy_vision(
    eco_innovation_by_bioeconomy_visions, codes_df
)

sns.move_legend(p, "lower center", ncol=5)
# TODO: perhaps a per cent display by group size would be interesting too?
# TODO: order y by label groups, color bars

# def barplot_eco_innovation_by_bioeconomy_vision(data):

```

@@

{{< pagebreak >}}
# References {.unnumbered}

@@

```{python}
# | label: tbl-full-codes
# | tbl-cap: "List of Eco-Innovation Labels"
format_table(codes_df.iloc[:5, :])

```

>>>>

---
title: Quantifying Directionality and Innovation Output in Sweden's Transition to a Forest-Based Bioeconomy
subtitle: Draft
author:
  name: Jonas Kreutzer
  affiliation: 'Department of Economic History, Lund University'
date: last-modified
abstract: |
  The transition to a bioeconomy is high on the agenda for research, industry and policy. Like other sustainability transition, the bioeconomy transition is characterized by contestation over desired outcomes and transition pathways. While shared visions are seen as pivotal in guiding transitions, the bioeconomy discourse is marked by the contention of three competing visions. The Swedish forest-based bioeconomy is well established as a case study in this arena. Yet, knowledge about innovation produced across the entire system is scarce. How innovative is this new economy, especially considering the central role of innovation in transition discourse? And what does commercialized innovation suggest for the directionality of the bioeconomy transition? The use of detailed innovation data from a Swedish literature-based innovation output database, SWINNO, allows to capture the actual innovation output of the Swedish forest-based bioeconomy innovation system. Qualitative coding of the source material links innovation from 1970 to 2021 with visions of a bioresource, biotechnology and / or bioecology bioeconomy. Findings reveal a notable decrease in the innovation output of the forest-based bioeconomy relative to Sweden's overall commercialized innovation. Notably, innovations associated with a bioresource vision surged during the energy crises of the 1970s, pre-dating their formalization in strategy documents, suggesting a significant role for path dependency in shaping shared visions. Concurrently, the presence of innovations aligning with multiple visions indicates an open opportunity space for future development across various directions. This research underscores the dynamic interplay between historical contexts, shared visions, and innovation trajectories within the bioeconomy transition, and questions the extent to which visions can provide novel directionality to transitions.

csl: 'https://zotero.org/styles/apa?source=1'
bibliography:
  - swinno-shades-green.bib
link-citations: true
link-bibliography: true
colorlinks: true
number-sections: true
execute:
  warning: false
  error: false
  echo: false
format:
  pdf:
    papersize: a4
    linestretch: 1.25
    mainfont: "Adobe Garamond Pro"
    sansfont: "Adobe Garamond Pro Semibold"
    include-in-header: 
        - text: \usepackage[a4paper, total={16.5cm,24.7cm}, top=2.5cm, ignorehead, centering, footskip=\footskip+4mm]{geometry}
        - text: \raggedbottom 
jupyter: python3
---

@@

import pandas as pd
import numpy as np
import scipy.stats as stats
from pyhere import here
from IPython.display import display, Markdown
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import seaborn as sns
from tabulate import tabulate

from src.figures import *
from src.utils import get_project_root, format_table
from src.data_handling import database

ROOT = get_project_root()
raw_data = here("data/raw")
modified_data = here("data/modified")
conn = database.connect_swinno_db()

@@

sns.set_theme(style="white")
```

```{python}
#| label: Load Data
raw_data = here("data/raw")
bioeconomy = pd.read_csv(raw_data / "id_year_bioeconomy.csv")
bioeconomy = bioeconomy.loc[bioeconomy["bioeconomy"] == 1]

swinno["bioeconomy"] = (
    swinno["sinno_id"].astype(str).isin(bioeconomy["sinno_id"].astype(str).unique())
)

@@

codes_df = codes_df.set_index("Code")

eco_innovations = database.query_eco_innovations(conn)
eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()

bioeconomy_visions = pd.read_csv(raw_data / "bioeconomy_visions.csv")
```

@@

<!-- Paragraphs 1-2: Motivation. After reading these paragraphs a reader in any field of economics should believe that if you answer your research question your paper will make an important contribution. -->
At the heart of sustainability transitions lies the question how societal functions can be fulfilled within ecological boundaries [@kohler2019AgendaSustainabilityTransitions].
Societal functions are provided through socio-technical systems, which consist of the interplay between the production, distribution, and use of technology [@geels2004SectoralSystemsInnovation].
Transitions are complex, long-term processes of shifts between different systems, involving diverse actors who have to chose between multiple, open-ended pathways, uncertain of their technological, social and political success [@kohler2019AgendaSustainabilityTransitions].
Further, success in the outcome and chosen pathway of sustainability transitions is contested based on the values, norms, and goals of  different actors and social groups.
Hence, innovation cannot be seen as a scalar, but must rather be understood as a vector, as possessing and providing direction to change rather than being simply progress forward [@stirling2011PluralisingProgressIntegrative].
@andersson2021OutcomesDirectionalityMorphology term these two properties _positive_ and _normative_ directionality, respectively.
Positive directionality of innovation arises from the fact that technological change is path-dependent [@arthur1989CompetingTechnologiesIncreasing, @david1985ClioEconomicsQWERTY], while normative directionality is rooted in the assumption that without public policy, private actors lack intrinsic incentives to pursue sustainability [@kohler2019AgendaSustainabilityTransitions].

A sustainability transition which has been gaining relevance for researchers, policy makers, and producers alike is the bioeconomy.
Broadly speaking, it aims to radically shift economic systems from relying on fossil fuels for energy and material resource, to an economic system built on biobased materials, circular and renewable value chains and higher environmental integrity through innovation[@bugge2016WhatBioeconomyReview].
As one of the main sources for biomass in a bioeconomy [@priefer2017PathwaysShapeBioeconomy], forests have become a key arena for the bioeconomy transition, highlighting tradeoffs between material provision, carbon storage and biodiversity [@hogbom2021TrilemmaNordicBaltic].
Hence, underneath the bioeconomy umbrella, the concept and viable transition pathways are heavily contested [@allain2022BioeconomicsVsBioeconomy].
Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition; the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview, @vivien2019HijackingBioeconomy].
The material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
How innovative is this new economy, especially considering the central role of innovation in transition discourse? And what does commercialized innovation suggest for the positive directionality of the bioeconomy transition?

<!-- @fischer2020SwedishForestsBioeconomy also note the consensus builidng of Swedish bioeconomy discourse marginalizes non-mainstream opinions. I can show that this is rooted, at least in part, in the path-dependence on technology output. -->

<!-- • Paragraphs 3-4: Challenges. These paragraphs explain why your research question has not already been answered, i.e., what are the central challenges a researcher must tackle to answer this question. -->
Rigorously operationalizing directionality is challening due to a plurality of possible development paths [@stirling2011PluralisingProgressIntegrative].
Stirling argues that innovation is better understood as a vector of possibilities rather than a scalar decision between clearly defined alternatives.
<!-- Lastly, Stirling (2011, 2009, 2008) is a clear voice pointing out a slightly different implication of directionality. Arguably, he introduced the concept to the field of sustainability transitions research, as part of broader critique of one-dimensional growth ideology. In line with a multidimensional view on sociotechnical change, he stresses that innovation is better viewed as “a vector, rather than just a scalar quantity” (Stirling, 2011, p. 84, 2008, p. 263). But he goes further and makes a passionate and eloquent call for a more pluralistic perspective on technological innovation.

@andersson2021OutcomesDirectionalityMorphology 111 -->

<!-- 1 The notion of directionality was introduced by Stirling (2008, 2009) as part of his call for opening up the process of technical change for alternative options. He builds on the broader idea that socio-technical change has a direction, choices are made between directions and actors gradually become blind to alternatives, which is a central tenet of much of the innovation studies literature. Our notion of directionality also draws on this idea (see also Weber and Rohracher, 2012: 1042–1043). @schot2018DeepTransitionsEmergence -->

@@

```{python}
#| label: Total SWINNO Innovation
#| output: asis

@@

| Corresponding Vision | Source article explicitly mentions that innovation:                                                   |
|----------------------|-------------------------------------------------------------------------------------------------------|
| Biotechnology        | Develops new, or modifies existing, organism, biological system, or process                            |
|                      | Uses cellular or molecular process to solve problem or in production process                          |
|                      | Applies biological agent in scientific or engineering based principle to produce product or service   |
|                      | Based on molecular biology, biochemistry, cell biology, embryology, genetics or microbiology research |
|                      | Provides method for research in the above areas                                                       |
| Bioresource          | Biomass used to substitute other input                                                                |
|                      | Refines biomass through mechanical, chemical or other process                                         |
|                      | Converts and upgrades bioresource                                                                     |
|                      | Optimises land use                                                                                    |
|                      | Includes degraded land in production of biofuels                                                      |
|                      | Intensifies use of biomass, including from waste                                                      |
| Bioecology           | Prioritizes biodiversity, conservation, ecosystem or soil health                                      |
|                      | Integrates production systems                                                                         |
|                      | Develops high quality product with territorial identity                                               |
|                      | Reuses or recycles waste                                                                              |
|                      | States compatibility with biosphere as motivation or aim                                              |
|                      | States need for social or ecological counter-expertise relative to other socio-technical innovations  |
|                      | Promotes or uses a circular production mode                                                           |
|                      | Promotes or uses a self-sustained production mode                                                     |

: Decision matrix used to classify innovation. {#tbl-decision-matrix tbl-colwidths="[20,80]"}

@@

### Eco-Innovation Types {#sec-eco_innovation_types}

@@

```{python}
#| label: Number Bioeconomy Innovation
#| output: asis

sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
```

@@

```{python}
#| label: Calculate Bioeconomy Share

@@

bioeconomy_share = calculate_bioeconomy_share(swinno)
```

```{python}
#| label: Bioeconomy Trend Description
#| output: asis
share_max = bioeconomy_share["bioeconomy_share"].max()
share_max_year = bioeconomy_share.loc[
    bioeconomy_share["bioeconomy_share"] == share_max, "year"
].values[0]
share_last_year = bioeconomy_share["year"].max()
share_latest = bioeconomy_share.loc[
    bioeconomy_share["year"] == share_last_year, "bioeconomy_share"
].values[0]
```

![a. A total number of innovation registered in SWINNO database. b. Forest based bioeconomy innovations registered in SWINNO database. c. Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database. Shown are the 5 year moving averages for each time series.](results/figures/innovation_trends.pdf){#fig-innovation-trends}

@@

```{python}
#| label: Uncertain Visions

@@

total_certain_bioeconomy_visions = (
    count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
    .agg("sum", numeric_only=True)
    .sort_values("count", ascending=False)
    .reset_index()
)
```

```{python}
#| label: Number of Uncertains
#| output: asis

n_uncertain_bioeconomy_visions = uncertain_bioeconomy_visions["sinno_id"].nunique()
```

For the following part, innovations which do not relate to the forest are excluded. Excluded are also the `{python} n_uncertain_bioeconomy_visions` innovations whose source text did not allow labeling without significant uncertaintiy. These innovations are planned to undergo a revision by subject matter experts at a later point.

```{python}
#| label: Number of Results
#| output: asis

@@

display(
    Markdown(
        f"This leaves {n_certain_bioeconomy_visions} innovations have been classified for the period between {certain_bioeconomy_visions_min_year} and {certain_bioeconomy_visions_max_year}. With {n_most_bioeconomy_vision} innovations, the {total_certain_bioeconomy_visions.iloc[0, 0]} has the most associated innovations. Next the {total_certain_bioeconomy_visions.iloc[1, 0]} was associated with {total_certain_bioeconomy_visions.iloc[1,1]}, while {total_certain_bioeconomy_visions.iloc[2,1]} and  {total_certain_bioeconomy_visions.iloc[3,1]} innovation were classified as {total_certain_bioeconomy_visions.iloc[2,0]} and {total_certain_bioeconomy_visions.iloc[3,0]}, respectively. Note that the sum by category will not yield the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision."
    )
)
```

@@

```{python}
#| label: Count Recycling Innovation

@@

```{python}
#| label: tbl-count_bioeconomy_visions
#| tbl-cap: Count of Bioeconomy Visions
#| output: asis
#| include: false

@@

![Counts of innovation classification by Bioeconomy Vision Category. Innovations can be classified into multiple visions simulatenously. The sum of all vision counts does not match the sum of all bioeconomy innovations.](results/figures/vision_count.pdf){#fig-bioeconomy-trends}

@@

Within the bioresource vision, two technological development blocks provide the basis for a substantial share of innovations: more efficient harvesting methods through technology and more efficient utilization of timber downstream.
Write here about the pendelsag, about the forwarders, about the extent to which sustainability was incorporated but only to the extent that it allowed more profitable harvesting of wood.

The role that the 1970s played in the development of core technologies for the bioresource vision can be illustrated in the fact that 7 of the 9 forwarder innovations recorded in SWINNO were commercialized between 1974 and 1982.
Forwarders are machines used to retrieve felled timber from forest plots before loading them on heavier timber transport options.
Forwarders replaced forest tractors or unmechanized timber retrival methods such as horse pulled sledges.
@nordfjell2019TechnicalDevelopmentForwarders show that although they have become heavier, their core design has not changed dramatically.

Perhaps a little bit about the uncertainty and promise that it would be better / not harmful to forest soil.

```{python}
#| label: forwarder
#| include: false
forwarder_innovation = pd.read_sql(
    "SELECT sinno_id, year_of_commercialization from innovation where description_in_swedish LIKE '%skotare%'",
    conn,
)

forwarder_innovation
```

Within bioecology, few innovations aim to decommodify value chains.
Dominated by technological development block to green the pollution of pulp and paper industry.

In biotechnology, attempts to form a development block around new material such as tencel flaunder.

@@

@@

eco_innovation_merged_bv["innovation_type_group"] = np.where(
    eco_innovation_merged_bv["innovation_type_code"].str[0] == "1",
    "Product",
    np.where(
        eco_innovation_merged_bv["innovation_type_code"].str[0] == "2",
        "Process",
        np.where(
            eco_innovation_merged_bv["innovation_type_code"].str[0] == "3",
            "Organizational",
            np.where(
                eco_innovation_merged_bv["innovation_type_code"].str[0] == "4",
                "Marketing",
                "No Eco-Innovation",
            ),
        ),
    ),
)

@@

```{python}
# | label: chi-square-vision-eco
other_sustainability = {501, 701, 108, 109, 111}
eco_innovation_filtered = eco_innovation_merged_bv[
    ~(
        eco_innovation_merged_bv["innovation_type_code"]
        .astype(int)
        .isin(other_sustainability)
    )
]
# Create the groups for the cross tab
eco_innovation_filtered["eco_innovation"] = (
    eco_innovation_filtered["innovation_type_group"] != "No Eco-Innovation"
)
eco_innovation_filtered["vision_neutral"] = (
    eco_innovation_filtered["bioeconomy_vision"] == "Vision Neutral"
)

# Calculate the chi squared test
vision_eco_cross_tab = (
    eco_innovation_filtered.drop_duplicates(subset="sinno_id")
    .groupby(["vision_neutral", "eco_innovation"])
    .agg(count=("sinno_id", "nunique"))
    .unstack()
)
chi2_stat, p_val, dof, expected = stats.chi2_contingency(vision_eco_cross_tab)
```

A noteworthy feature of innovation belonging to No Bioresource Vision is that this category has by far the most Non-Eco innovations.
The chi-squared test of independence reveals a statistically significant association between an innovation being aligned with a vision and it being an eco-innovation ($\chi^2 =$ `{python} round(chi2_stat, 3)` df = `{python} dof`, p < 0.05).
For the sustainability transition of the forest bioeconomy this could imply that having some Vision of a bioeconomy creates more eco-innovation than pursuing other innovation paths.

```{python}
# | label: tbl-cross-table-eco-visions
# | tbl-cap: Cross Table of Unique Innovations Classified as Vision Neutral And Eco-Innovation
# Headers
headers = ["Non-Eco-Innovation", "Eco-Innovation"]
index = ["Bioeconomy Vision", "Vision Neutral"]

# Generate the table
cross_table = tabulate(
    vision_eco_cross_tab.values, headers=headers, showindex=index, tablefmt="pipe"
)

display(Markdown(cross_table))
```

```{python}
#| label: fig-eco-innovation-bioeconomy-vision
#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
p = plot_eco_innovation_by_bioeconomy_vision(
    eco_innovation_by_bioeconomy_visions, codes_df
)

sns.move_legend(p, "lower center", ncol=5)
# TODO: perhaps a per cent display by group size would be interesting too?
```

@@

{{< pagebreak >}}

# References {.unnumbered}

@@

```{python}
#| label: tbl-full-codes
#| tbl-cap: List of Eco-Innovation Labels
format_table(codes_df.iloc[:5, :])
```
